---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(knitr)
require(ggplot2)
require(lubridate)
require(RColorBrewer)
require(dplyr)
library(plyr)
require(ggthemes)
library(scales)
loadData <- function(dataURL="", destF="default.csv", method = NULL){
  if(!file.exists(destF)){
            temp <- tempfile()
            download.file(dataURL, temp, method = method)
            unzip(temp, destF)
            unlink(temp)
        }else{
            message("Data already downloaded.")
        }
}

dataURL <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
loadData(dataURL, "activity.csv")
#
active <- read.csv("activity.csv")
active$date<-as.Date(active$date)
head(active)
```


## What is mean total number of steps taken per day?
```{r}
total.steps <- tapply(active$steps, active$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
meanTotalSteps <- mean(total.steps, na.rm=TRUE)
#cat (" The mean  total number of steps taken per day is :", meanTotalSteps  )

```
The mean total number of steps per day is `r meanTotalSteps` 



## What is the average daily activity pattern?
```{r}
spi <- ddply(active, .(interval), summarize, steps = mean(steps, na.rm=TRUE))
max.spi <- max(spi$steps)
max.int <- spi[spi$steps==max(max.spi),1]
```
**Plot of the average number of steps taken each interval**
```{r}
plot(spi$interval,spi$steps, ylab="Average number of steps per interval",xlab="Interval", main=NULL, type="l", lwd=2, col="green")
abline(h=max.spi, col = c("red"))
abline(v=max.int, col = c("black"))
```
The maximum number of steps per interval in average is `r max.spi` and corresponds to the interval `r max.int`.

## Imputing missing values
```{r}
missingValues <- dim(active[is.na(active$steps),])[1]
```
There are  `r missingValues` with missing values

**all missing values are filled in with the average value of the 5-minute intervals**
```{r}
active[is.na(active$steps), "steps"] <- mean(active$steps, na.rm=T)
missing <- is.na(active$steps)
# How many missing
table(missing)
```

**Histogram with  total number  of steps  taken each day and calculate and respor the mean  an median  total number of steps aken per day**
```{r}

spd1 <- ddply(active, .(date), summarize, steps = sum(steps))
mean.spd1 <- mean(spd1$steps, na.rm = TRUE)
median.spd1 <- median(spd1$steps, na.rm = TRUE)
hist(active$steps, xlab="Total number of steps per day", main=NULL, col="light green")
abline(v=mean.spd1, col = c("red"))
abline(v=median.spd1, col = c("green"), lty = 2)

```




## Are there differences in activity patterns between weekdays and weekends?
